Parsing Postfix log files

### 2007-06-30 doc/logparser.tex
Documented the action that handles watchdog timeouts. Expanded the mails covered section, listing the warnings from test-some. Removed limitations and improvements which no longer apply.

@@

@@

Coverage of mails is much more difficult to determine accurately.  The
parser can dump its state tables in a human readable form; examining these
tables with reference to the log files is the best way to detect mails
which were not handled properly (all the complications discussed in
section~\ref{additional complications} were detected in this way).  The
parser issues warnings when it detects any errors, some of which may alert
the user to a problem, e.g.\ when a queueid is reused before the previous
mail is fully dealt with, when a queueid/pid is not found, or when there
are problems tracking a child mail (see
section~\ref{tracking-re-injected-mail}).  There should be few or no
warnings (warnings about mail where some of the log lines are not in the
logfile can be ignored), and the state table should only contain entries
for mails which had yet to be delivered when the logfiles ended.

At the time of writing the parser is being tested with 93 log files.  The
state table contains entries for mails not yet delivered at the end of the

@@

    \item The program doesn't cope with mails which span log files:
        currently they are reported as unfinished mails at the end of the
        first run and are complained about when encountered during the next
        run because they don't have a source.

@@

@@

    \item Save state at the end of each run; load it at the start of the
        next run.  This should allow the program to cope with mails
        spanning log files.  This might also require purging old mails once
        they've been in the state tables for too long, though the presence
        of very old mails indicates a bug in either the implementation or
        the specification.

>>>>

    \item [SMTPD\_WATCHDOG] Smtpds have a watchdog timer to deal with
        unusual situations --- after five hours the timer will expire and
        the smtpd will exit.  This occurs very infrequently, as there are
        many other timeouts which should occur in the intervening hours:
        DNS timeouts, timeouts reading data, etc.  The active connection
        for that smtpd must be committed and deleted from the state tables.

@@

Coverage of mails is much more difficult to determine accurately than
coverage of lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly (all the
complications discussed in section~\ref{additional complications} were
diagnosed and sometimes detected in this way).  The parser issues warnings
when it detects any errors, some of which may alert the user to a problem,
e.g.\ when a queueid is reused before the previous mail is fully dealt
with, when a queueid or pid is not found,\footnote{There will often be a
few warnings about missing queueids or pids in the first few hundred or
thousand log lines because the earlier log lines are in the previous log
file; loading the saved state from the previous log file will solve this
problem} or when there are problems tracking a child mail (see
section~\ref{tracking-re-injected-mail}).  There should be few or no
warnings, and the state table should only contain entries for mails which
had yet to be delivered when the logfiles ended.

At the time of writing the parser is being tested with 93 log files.  There
are 25 warnings produced, but because the parser errs on the side of
producing more warnings rather than fewer those 25 warnings represent only
8 problems:

    \item 3 connections (5 warnings) started before the first log file, so
        their initial log entries are missing.

    \item 5 mails (20 warnings) have out of order log entries, leading to
        warnings about missing data and queueids.  The two log files where
        the out of order log lines occur are more than an order of
        magnitude larger than normal, due to the extremely large volume of
        mail handled by the server on those days; process scheduling
        vagaries and the undefined order of messages sent to syslog(8) may
        explain why the lines were out of order.

The first of these problems cannot be dealt with; although out of order log
lines (the second problem) are dealt with in other circumstances (see
sections~\ref{tracking-re-injected-mail}, \ref{discarding-cleanup-lines}
and \ref{pickup-logging-after-cleanup}), dealing with these lines would
introduce a large amount of complication (deferred warnings, deferred
committal, ambiguity) for a problem which occurs only four times in
60721709 log lines.

@@

The
state table contains entries for mails not yet delivered at the end of the

@@

@@

        XXX EXPAND ON THIS

@@
